Prediction of PIK3CA mutation with gene expression

### 2020-04-22 plosOne/plosOne.Rmd
discussion RAS

---
title: Prediction of PIK3CA mutation with gene expression
author:
  - name: Jun Kang
    email: jkang.alien@gmail.com
    affiliation: Department of Hospital Pathology, Seoul St. Mary’s Hospital, College of Medicine, The Catholic University of Korea, Seoul, South Korea
    corresponding: jkang.alien@gmail.com
abstract: |
bibliography: mybibfile.bib
output: rticles::plos_article
csl: plos.csl
---

@@

A list

@@

The problem

- PIK3CA mutation in selecting drug

@@

RNAseq can be widely used. The mutation status directly prediction

Previous study @way_2018_machine_cellreports

prediction of aberant activation of a certain pathway vs a specific mutation
The proposed solution

@@

To narrow down potential predictors, Genes with a large the median absolute deviation (more than third-quartiles) were selected. 5000 out of 20502 genes were included in the modeling process.

Yeo-Johnson transformation was done to correct skewness. Centering and scaling were done. All preprocessing was done using recipe r package. [@kuhnRecipesPreprocessingTools2020]

Penalized logistic regression was applied to prediction modeling. 10-fold cross-validation with targe variable stratification was done over the hyperparameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75}. Model performance was evealuated with the area under receiver operating characteristic (ROC) and area under the ROC (AUROC) and precision recall (PR) and area under curve (AUPR). The modelin process was done with tidymodels rpackage.  [@kuhnTidymodelsEasilyInstall2020]

@@

## selecting model

@@

## performance of each cancer type  
Because tht prevalence of PIK3CA mutation is vary across the cancer type, the performance of each cancer type was investigated. The AUROC and AUPR were positively correlated between train set and test set in cancer type subanalysis. (Figure 1B) The AUPR was high in cancer type with high PIK3CA mutation rate such as colon, brest, Uterus cancer types. The AUROC did not correlated with PIK3CA mutation rate of each cancer types. (Figure 1C)

@@

@@

Main message answers the question and main supporting evidence

Critical assessment opinions on 
- any shortcomings in study design

@@

@@

Case imbalance

- flaws in analysis

- validity of assumption

Comparison with other studies where inconsistencies are discussed

PIK3CA prediction model performed better than RAS activation prediction model in terms of AUROC. AUPR and the PIK3CA mutation rate interpretation. @davis_2006_relationship_proc23rdintconfmachlearn

Conclusions comments on possible biological or clinical implications and suggestions for further research.  
Evaluate the results - not the authors

@@

Figure 1. Summary of modeling results

@@

Figure 2. Coefficients of model

@@

S1 Appendix.  
S2 Figure.  
S1 Table.  
S2 Table.  
S3 Table.

>>>>

---
title: Prediction of PIK3CA mutation with gene expression
author:
  - name: Jun Kang
    email: jkang.alien@gmail.com
    affiliation: Department of Hospital Pathology, Seoul St. Mary’s Hospital, College of Medicine, The Catholic University of Korea, Seoul, South Korea
    corresponding: jkang.alien@gmail.com
abstract: |
bibliography: [mybibfile.bib, pik3ca.bib] 
indent: true
output: rticles::plos_article
csl: plos.csl
---

@@

@@

- The problem  
  * PIK3CA mutation in selecting drug

@@

- RNAseq can be widely used. The mutation status directly prediction

- Previous study @way_2018_machine_cellreports

- Prediction of aberant activation of a certain pathway vs a specific mutation

- The proposed solution

@@

To narrow down potential predictors, Genes with a large the median absolute deviation (more than third-quartiles) were selected. 5000 out of 20502 genes were included in the modeling process. Yeo-Johnson transformation was done to correct skewness. Centering and scaling were done. All preprocessing was done using recipe r package. [@kuhnRecipesPreprocessingTools2020] Penalized logistic regression was applied to prediction modeling. 10-fold cross-validation with targe variable stratification was done over the hyperparameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75}. Model performance was evealuated with the area under receiver operating characteristic (ROC) and area under the ROC (AUROC) and precision recall (PR) and area under curve (AUPR). The modelin process was done with tidymodels rpackage.  [@kuhnTidymodelsEasilyInstall2020]

@@

## Selecting model

@@

## Performance of each cancer type  
Because tht prevalence of PIK3CA mutation is vary across the cancer type, the performance of each cancer type was investigated. The AUROC and AUPR were positively correlated between train set and test set in cancer type subanalysis. (Figure 1B) The AUPR was high in cancer type with high PIK3CA mutation rate such as colon, brest, Uterus cancer types. The AUROC did not correlated with PIK3CA mutation rate of each cancer types. (Figure 1C)

@@

Figure 2 shows top 30 important predictors. IGF1R mRNA expression was the strongest negative predictor and PTEN was the strongest positive predictor. Both genes and PIK3CA are key players in tyrosin kanase pathway. The cancer type was important predictors. Some cancer types including uterine carcinosarcoma (UCS), bladder urothelial carcinoma (BLCA), pancreatic adenocarcinoma (PAAD), lymphoid neoplasm diffuse large B-cell lymphoma (DLBC) are strongest predictors.

@@

- Main message answers the question and main supporting evidence

- Critical assessment opinions on any shortcomings in study design

@@

In this study, the prevalence of PIK3CA mutation was 11%. The low prevalence of PIK3CA results in imbalaced dataset which makes the prediction difficult. The AUROC is optimistic in terms of performance. The AUPR is more informative than AUROC on imbalaced datasets. [@saitoPrecisionRecallPlotMore2015]

@@

  * Case imbalance  
  * flaws in analysis  
  * validity of assumption

- Comparison with other studies where inconsistencies are discussed  
- Evaluate the results - not the authors

Our PIK3CA prediction model performed better than RAS activation prediction model of previous study in terms of both AUROC (0.84 vs 0.75) and AUPR (0.39 vs 0.24) on testset prediction. Our testset is corresponding to the samples initially filtered from training. The target variable of our study is more specific than the previous study. The specific important mutations can effect stronger downsteam gene expression than the broad events pathway activation. The previous study might be more difficult prediction problum.

- Conclusions comments on possible biological or clinical implications and suggestions for further research.

@@

- Figure 1. Summary of modeling results

@@

- Figure 2. Coefficients of model

@@

- S1 Appendix.  
- S2 Figure.  
- S1 Table.  
- S2 Table.  
- S3 Table.

### 2020-04-22 plosOne/plosOne.tex
discussion RAS

A list

\begin{itemize}
\tightlist
\item
  Brevity
\item
  Logic and clarity
\item
  Clean typing
\end{itemize}

The problem

\begin{itemize}
\tightlist
\item
  PIK3CA mutation in selecting drug
\end{itemize}

@@

RNAseq can be widely used. The mutation status directly prediction

Previous study {[}2{]}

prediction of aberant activation of a certain pathway vs a specific
mutation The proposed solution

@@

To narrow down potential predictors, Genes with a large the median
absolute deviation (more than third-quartiles) were selected. 5000 out
of 20502 genes were included in the modeling process.

Yeo-Johnson transformation was done to correct skewness. Centering and
scaling were done. All preprocessing was done using recipe r package.
{[}5{]}

Penalized logistic regression was applied to prediction modeling.
10-fold cross-validation with targe variable stratification was done
over the hyperparameter grid: \(\lambda\) \{\(10^{-5}\),
\(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\), \(10^{0}\)\},
\(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Model performance was evealuated
with the area under receiver operating characteristic (ROC) and area
under the ROC (AUROC) and precision recall (PR) and area under curve
(AUPR). The modelin process was done with tidymodels rpackage. {[}6{]}

@@

\hypertarget{selecting-model}{%
\subsection{selecting model}\label{selecting-model}}

@@

\hypertarget{performance-of-each-cancer-type}{%
\subsection{performance of each cancer
type}\label{performance-of-each-cancer-type}}

@@

Main message answers the question and main supporting evidence

Critical assessment opinions on - any shortcomings in study design

@@

\begin{itemize}
\tightlist
\item
  limitations in methods
\end{itemize}

Case imbalance

\begin{itemize}
\item
  flaws in analysis
\item
  validity of assumption
\end{itemize}

Comparison with other studies where inconsistencies are discussed

@@

Conclusions comments on possible biological or clinical implications and
suggestions for further research.\\
Evaluate the results - not the authors

@@

Figure 1. Summary of modeling results

@@

Figure 2. Coefficients of model

@@

S1 Appendix.\\
S2 Figure.\\
S1 Table.\\
S2 Table.\\
S3 Table.

@@

\leavevmode\hypertarget{ref-davis_2006_relationship_proc23rdintconfmachlearn}{}%
7. Davis J, Goadrich M. The relationship between Precision-Recall and
ROC curves. Proceedings of the 23rd international conference on Machine
learning. Pittsburgh, Pennsylvania, USA: Association for Computing
Machinery; 2006. pp. 233--240.
doi:\href{https://doi.org/10.1145/1143844.1143874}{10.1145/1143844.1143874}

>>>>

\begin{itemize}
\item
  Brevity
\item
  Logic and clarity
\item
  Clean typing
\item
  The problem

  \begin{itemize}
  \tightlist
  \item
    PIK3CA mutation in selecting drug
  \end{itemize}
\end{itemize}

@@

\begin{itemize}
\item
  RNAseq can be widely used. The mutation status directly prediction
\item
  Previous study {[}2{]}
\item
  Prediction of aberant activation of a certain pathway vs a specific
  mutation
\item
  The proposed solution
\end{itemize}

@@

To narrow down potential predictors, Genes with a large the median
absolute deviation (more than third-quartiles) were selected. 5000 out
of 20502 genes were included in the modeling process. Yeo-Johnson
transformation was done to correct skewness. Centering and scaling were
done. All preprocessing was done using recipe r package. {[}5{]}
Penalized logistic regression was applied to prediction modeling.
10-fold cross-validation with targe variable stratification was done
over the hyperparameter grid: \(\lambda\) \{\(10^{-5}\),
\(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\), \(10^{0}\)\},
\(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Model performance was evealuated
with the area under receiver operating characteristic (ROC) and area
under the ROC (AUROC) and precision recall (PR) and area under curve
(AUPR). The modelin process was done with tidymodels rpackage. {[}6{]}

@@

\hypertarget{selecting-model}{%
\subsection{Selecting model}\label{selecting-model}}

@@

\hypertarget{performance-of-each-cancer-type}{%
\subsection{Performance of each cancer
type}\label{performance-of-each-cancer-type}}

@@

\begin{itemize}
\item
  Main message answers the question and main supporting evidence
\item
  Critical assessment opinions on any shortcomings in study design
\end{itemize}

@@

In this study, the prevalence of PIK3CA mutation was 11\%. The low
prevalence of PIK3CA results in imbalaced dataset which makes the
prediction difficult. The AUROC is optimistic in terms of performance.
The AUPR is more informative than AUROC on imbalaced datasets. {[}7{]}

\begin{itemize}
\item
  limitations in methods

  \begin{itemize}
  \tightlist
  \item
    Case imbalance\\
  \item
    flaws in analysis\\
  \item
    validity of assumption
  \end{itemize}
\item
  Comparison with other studies where inconsistencies are discussed\\
\item
  Evaluate the results - not the authors
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Conclusions comments on possible biological or clinical implications
  and suggestions for further research.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Figure 1. Summary of modeling results
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Figure 2. Coefficients of model
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  S1 Appendix.\\
\item
  S2 Figure.\\
\item
  S1 Table.\\
\item
  S2 Table.\\
\item
  S3 Table.
\end{itemize}

@@

\leavevmode\hypertarget{ref-saitoPrecisionRecallPlotMore2015}{}%
7. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
than the ROC Plot When Evaluating Binary Classifiers on Imbalanced
Datasets. PLOS ONE. Public Library of Science; 10: e0118432.
doi:\href{https://doi.org/10.1371/journal.pone.0118432}{10.1371/journal.pone.0118432}
